---
title: "Using the CCEB High Performance Computing Cluster: Commands and References"
author: "Alessandra Valcarcel"
date: 2019-09-03T21:01:01-06:00
categories: ["Computing"]
tags: ["Cluster", "CCEB", "LSF", "IBM"]
output: 
  bookdown::html_document2:
    number_sections: false
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(out.width = '80%',
                      fig.align = 'center', 
                      echo=TRUE,
                      eval = FALSE, 
                      warning=FALSE, 
                      message=FALSE)
```

# Introduction

This is the tenth and final blog post in a series of articles about using the CCEB cluster. An overview of the series is available [here](www.alessandravalcarcel.com/blog/2019-04-13-clusterintro/). This post includes all the commands and references used throughout the series. 

# Commands

The commands presented below are sometimes generalizations of what was presented throughout the series.

## Command Line

```{bash}
# Change the working directory
cd 
# List folders/files in a directory
ls 
# Report working directory
pwd 
# Move files or directories
mv 
# Copy files or directories
cp
# Summons the manual
man 
# Create a directory
mkdir 
# Removes a directory
rmdir 
# Removes files or a directory
rm 
# Clear the screen/command board in Terminal
clear 
# Uses all files with the pattern indicated in the command applied
*pattern* 
# Creates an empty text file
touch 
```

## Log On

```{bash, echo = TRUE, eval = FALSE}
ssh -X pennkey@scisub
```

## Screen

```{bash, eval = FALSE}
# Screen the session while on the scisub host
screen
# Check how many screens you have open 
screen -ls
# Resuming a screen
screen -r
```

## Interactive Session

```{bash, eval = FALSE}
# Full bsub for an interactive session
bsub -Is -q cceb_interactive -m "silver01" -n 8 -R  "span[hosts=1] rusage[mem=5000]" -M 10000 "bash"
```

```{r}
# R code to detect the number of cores requested by bsub
# Recall parallel::detectCores() does not report the proper number
Sys.getenv('LSB_DJOB_NUMPROC')
```


## Normal Session

```{bash}
bsub -q cceb_normal \
-J "jobname[1-n]%5" \
-m "silver02" \
-R "rusage[mem=5000]" \
-M 10000 \
-n 8 \
-o /project/taki3/amv/cluster/normal_session_examples/example_output.txt \
Rscript /project/taki3/amv/cluster/normal_session_examples/normal_example.R
```

```{r}
# Get the normal batch job index in R
Sys.getenv("LSB_JOBINDEX")
```

## Quality Control

```{bash}
# Show queues
bqueues -u alval
# List execution host information
bhosts
# Check a queues traffic
bjobs -u all | grep cceb
# List all your jobs (pending and running)
bjobs
# List all running jobs
bjobs -r
# List all pening jobs
bjobs -p
# List all suspended jobs
bjobs -s
# List summary information
bjobs -sum
# Estimate run time
bjobs -WF
# Check history of current jobs running
bhist
# Use the job name to peek
bpeek -J "jobname[38]"
# Use the job ID to peek
bpeek 2147852[38]
# Suspend or pause a job
bstop
# Resume a job
bresume
# Kill a job 
bkill
# Check live machine usage -- must be run on execution host (i.e. after bsub)
top
```

# References

[A very nice LSF cheat sheet](https://www.tu-ilmenau.de/fileadmin/media/unirz/Services/Struktureinheiten/Advanced_Computing/lsf6.2_quick_ref.pdf)

[Command line basic youtube video](https://www.youtube.com/watch?v=5XgBd6rjuDQ)

## Setup

[VPN](https://en.wikipedia.org/wiki/Virtual_private_network)

[Pulse Secure](https://www.pulsesecure.net/)

[Download Pulse Secure here](https://www.pulsesecure.net/trynow/client-download/)

[Generating `ssh` key-pairs to bypass password input after logging onto cluster](https://www.thegeekstuff.com/2008/11/3-steps-to-perform-ssh-login-without-password-using-ssh-keygen-ssh-copy-id)

[To continue a `bsub` on multiple lines](https://stackoverflow.com/questions/3871332/how-to-tell-bash-that-the-line-continues-on-the-next-line)

[`.bash_profile`](https://www.quora.com/What-is-bash_profile-and-what-is-its-use)

[How To Use Linux Screen](https://linuxize.com/post/how-to-use-linux-screen/)

[Screen manual](https://linux.die.net/man/1/screen)

[HPC computing in R](https://www.hpc.dtu.dk/?page_id=2723)

[`top`](https://www.tecmint.com/12-top-command-examples-in-linux/)

## Penn

[PMACs ticket](https://helpdesk.pmacs.upenn.edu/)

[HPC wiki](https://dbe.med.upenn.edu/secure/wiki/index.php?title=Linux_Grid_Help)

## IBM Hardware

[Introduction to IBM Platform LSF](https://www.ibm.com/support/knowledgecenter/en/SSETD4_9.1.2/lsf_foundations/lsf_introduction_to.html)

[IBM Hosts](https://www.ibm.com/support/knowledgecenter/en/SSETD4_9.1.3/lsf_users_guide/hosts_about.html)

[How that Node/Socket/Core thing works](https://blogs.technet.microsoft.com/windowshpc/2008/04/14/how-that-nodesocketcore-thing-works/)

[Clusters, jobs, and queues](https://www.ibm.com/support/knowledgecenter/en/SSETD4_9.1.2/lsf_users_guide/clusters_jobs_about.html)


## IBM Getting Started

[Getting started with IBM Spectrum LSF](https://www.ibm.com/support/knowledgecenter/en/SSWRJV_10.1.0/lsf_welcome/lsf_kc_get_started.html)

[IBM Platform LSF](https://www.ibm.com/support/knowledgecenter/en/SSETD4/product_welcome_platform_lsf.html)

[What is batch processing](https://www.ibm.com/support/knowledgecenter/zosbasics/com.ibm.zos.zconcepts/zconc_whatisbatch.htm)

[Control job execution with bstop, bresume, and bkill](https://www.ibm.com/support/knowledgecenter/en/SSWRJV_10.1.0/lsf_admin_foundations/control_job_exec.html)

## IBM Command Manuals

[`bsub`](https://www.ibm.com/support/knowledgecenter/en/SSETD4_9.1.2/lsf_command_ref/bsub.1.html)

[`bqueues`](https://www.ibm.com/support/knowledgecenter/en/SSETD4_9.1.3/lsf_command_ref/bqueues.1.html) - Display information about queues.

[`bhosts`](https://www.ibm.com/support/knowledgecenter/en/SSETD4_9.1.3/lsf_command_ref/bhosts.1.html) - Display information about execution hosts.

[`bjobs`](https://www.ibm.com/support/knowledgecenter/en/SSETD4_9.1.2/lsf_command_ref/bjobs.1.html) - Check job status .

[`bhist`](https://www.ibm.com/support/knowledgecenter/en/SSETD4_9.1.2/lsf_command_ref/bjobs.1.html) - Check job and task status history and summary.

[`bpeek`](https://www.ibm.com/support/knowledgecenter/en/SSETD4_9.1.3/lsf_command_ref/bpeek.1.html) - Peek into job to view output.

[`bstop`](https://www.ibm.com/support/knowledgecenter/en/SSETD4_9.1.2/lsf_command_ref/bstop.1.html) -   Suspend a job.

[`bresume`](https://www.ibm.com/support/knowledgecenter/en/SSETD4_9.1.2/lsf_command_ref/bresume.1.html) -  Resume a suspended job.

[`bkill`](https://www.ibm.com/support/knowledgecenter/en/SSETD4_9.1.3/lsf_command_ref/bkill.1.html) - Terminate a job.
